---
title: "DEP_Barron Lab_050821"
author: 'Author: Lai, Kei Onn (Barron lab) and Giuseppe D Agostino (Langley Lab)'
date: "8/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(xlsx)
library(DEP)
library(impute)
library(dplyr)
library(SummarizedExperiment)
library(MSnbase)
library(tibble)
library(imp4p)
library(bnstruct)
```
Import Empai Data, ensure all data(apart from protein Accession and Description) is numerical 
```{r}
setwd("C:/Users/Lim Peck Hui/Desktop/path/050821")
empai_results<-read.csv(file="TSPO_empai_original.csv",row.names = 1)
colnames(empai_results)
# Create coldata
dat<-empai_results

dat[,3:ncol(dat)] <- sapply(dat[,3:ncol(dat)],as.numeric)
colnames(dat)[1] <- "ID"

coldata <- data.frame("label" = colnames(dat)[3:ncol(dat)], 
"condition" = as.factor(c(rep("WT", 3), rep("APP", 3), rep("KO", 3))), "replicate" = as.factor(c(1:3, 1:3, 1:3)))
#have a look how coldata looks like
coldata
```

Creating summarized Experiment with coldata
Herein, samples in log2 format
```{r}

data_unique <- make_unique(dat, "ID", "Description")

dat_se <- make_se(data_unique, 3:11, coldata)

assay<- data.frame(assay(dat_se))
assay$ID<-row.names(assay)
write.table(assay,file="assay_datse.txt")
write.table(dat,file="originalempaiinput.txt")
```

Finding proteins present in all 3 WT samples but absent in all 3 KO samples and all 3 APP samples; we term these proteins as part of Differentially enriched proteins (DEP) between WT and APP

```{r}
WT_true_APPNA<-dat[!is.na(dat$X191020_TSPO_HB15WTMod.xlsx)&!is.na(dat$X200111_TSPO_A6WTMod.xlsx)&!is.na(dat$X200111_TSPO_B6WTMod.xlsx)&is.na(dat$X191020_TSPO_SF911APPMod.xlsx)&is.na(dat$X200111_TSPO_6APPF2Mod.xlsx)&is.na(dat$X200111_TSPO_6APPF6Mod.xlsx),]

#row 7 aka last row has value in one of KO col.
WT_true_APPNA<- WT_true_APPNA[1:6,]
```

Finding proteins present in all 3 APP samples but absent in all 3 KO samples and all 3 WT samples; we term these proteins as part of Differentially enriched proteins (DEP) between APP and WT


```{r}
APP_true_WTNA<-dat[is.na(dat$X191020_TSPO_HB15WTMod.xlsx)&is.na(dat$X200111_TSPO_A6WTMod.xlsx)&is.na(dat$X200111_TSPO_B6WTMod.xlsx)&!is.na(dat$X191020_TSPO_SF911APPMod.xlsx)&!is.na(dat$X200111_TSPO_6APPF2Mod.xlsx)&!is.na(dat$X200111_TSPO_6APPF6Mod.xlsx),]

APP_true_WTNA<-APP_true_WTNA[is.na(APP_true_WTNA$X191020_TSPO_HB1KOMod.xlsx)&is.na(APP_true_WTNA$X200111_TSPO_B5KOMod.xlsx)&is.na(APP_true_WTNA$X200111_TSPO_D8KOMod.xlsx),]

rownames(APP_true_WTNA)<-APP_true_WTNA$ID

rownames(WT_true_APPNA)<-WT_true_APPNA$ID

APProwidx<-which(rownames(assay) %in% rownames(APP_true_WTNA)) 
WTrowidx<-which(rownames(assay) %in% rownames(WT_true_APPNA))
```
Imputing Min Prob for the 10 Differentially enriched proteins (DEP) after subtracting KO background
```{r}
DEPimputation <- dat_se
MSnSet <- as(DEPimputation, "MSnSet")

#for WT true, APP NA to be imputed in APP columns 4:6
MSnSetimp_minprobAPP<- MSnbase::impute(MSnSet[, 4:6], method = "MinProb")
DEPMinProbAPP<-MSnbase::exprs(MSnSetimp_minprobAPP)
DEPMinProbAPP[WTrowidx,]

#for APP true, WT NA to be imputed in WT columns 1:3
MSnSetimp_minprobWT<- MSnbase::impute(MSnSet[, 1:3], method = "MinProb")
DEPMinProbWT<-MSnbase::exprs(MSnSetimp_minprobWT)

#we just need to impute the 10 DEP proteins into assay
#here assay that has been imputed is called DEPimputation
assay(DEPimputation)[WTrowidx,4:6]<- DEPMinProbAPP[WTrowidx,]
assay(DEPimputation)[APProwidx,1:3]<-DEPMinProbAPP[APProwidx,]
```

let’s compare assay(DEPimputation) -which is after DEP imputation and assay(dat_se)-which is before DEP imputation

```{r}
#This shows WT true APP NA ; but these APP NAs are before imputation
assay[WTrowidx,]

#This shows WT true APP NA ; but these APP NAs have been imputed
assay(DEPimputation)[WTrowidx,]

#This shows APP true WT NA ; but these WT NAs are before imputation
assay[APProwidx,]
#This shows APP true WT NA ; but these WT NAs have been imputed
assay(DEPimputation)[APProwidx,]
```
Filter the 80% presence in APP+ WT comb from DEPimputation
```{r}
#check that col 1:6 corresponds to WT1-3 and APP1-3 respectively
DEPimputation[,1:6]

#filtering for proteins 80% present in APP vs WT comb
data_filt <- filter_proteins(DEPimputation[,1:6], type="fraction",min=0.8)
dim(data_filt)

```
We have 335 hits after filtering for 80% presence in APPvs WT comb as compared to KO background; we now impute for KO columns for Min Prob

```{r}
dim(data_filt)
#we noticed KO columns not in data_filt, need to put them back
data_filt_assay<-data.frame(assay(data_filt))

data_filt_assay<-merge(data_filt_assay,assay(DEPimputation[,7:9]),by.x=0,by.y=0,all.x=TRUE)

head(data_filt_assay)
names(data_filt_assay)[names(data_filt_assay) == "Row.names"] <- "ID"

ID<-dat[,1:2]

data_filt_assay<-merge(data_filt_assay,ID,by.x="ID",by.y="ID",all.x=TRUE)

row.names(data_filt_assay)<-data_filt_assay$ID

filtprot.idx<-which(row.names(data_filt_assay)%in%row.names(assay))
dat_se_filt <- dat_se[filtprot.idx,]
head(data_filt_assay)
#dropping col 1 and 11 which corresponds to ID and description, as ID is already in row.names
head(data_filt_assay[,-c(1,11)])

data_filt_assay<-data_filt_assay[,-c(1,11)]

````

KO imputation by MinProb
````{r}

data_filt_assay<-merge(data_filt_assay,ID,by.x=0,by.y="ID",all.x=TRUE)
#make back summarized expt

names(data_filt_assay)[names(data_filt_assay) == "Row.names"] <- "ID"

dat_filt_unique <- make_unique(data_filt_assay, "ID", "Description")
dat_filt_unique_data<-dat_filt_unique[,c(2:10)]
rownames(dat_filt_unique_data)<-dat_filt_unique$name

colnames(dat[3:ncol(dat)])
originalcolnames<-colnames(dat[3:ncol(dat)])
names(dat_filt_unique_data)<-originalcolnames
dat_filt_unique_data<-dat_filt_unique_data %>%
  rownames_to_column("ID")

dat_filt_unique_data<- merge(dat_filt_unique_data,ID,by.x="ID",by.y="ID",all.x=TRUE)


dat_filt_unique <- make_unique(dat_filt_unique_data, "ID", "Description")

head(dat_filt_unique)
dat_filt_unique[,2:10]<-sapply(dat_filt_unique[,2:10],as.numeric)
#previously this dat_filt_unique was logged 2, we need to revert back before putting back to summarized expt again
dat_filt_unique[,2:10]<-2^(dat_filt_unique[,2:10])


dat_se_filt2 <- make_se(dat_filt_unique, 2:10, coldata)
head(assay(dat_se_filt2))
#for KO columns 7:9

assay(dat_se_filt2[,7:9])<-impute.MinProb(assay(dat_se_filt2[,7:9]))


```

Impute KNN method for APP+WT samples combined

```{r}
APPWTknnimp<-impute.knn(assay(dat_se_filt2[,1:6]),k=5)
assay(dat_se_filt2[,1:6])<-APPWTknnimp$data
```
Revert the log and export final imputed data 
```{r}
finalimputeddata<-2^(assay(dat_se_filt2))
write.xlsx(finalimputeddata, 
           file="before and afterfinalimputeddata.xlsx",sheetName="finalimputeddata",
           append=TRUE)
write.xlsx(dat, "before and afterfinalimputeddata.xlsx", sheetName="Originalempai", append=TRUE)
```

Output of this saved xlsx sheet will be input of Differential Enrichment Workflow RMarkdown
(2.IPTSPO_DifferentiallyEnrichedWorkflowFinal_Github.Rmd)
